\startappendix{Interview Questions - Instructor Perspective}

This appendix lists the interview questions we used in the study highlighted in Chapter 4. The first section lists the starting set of questions we asked the educators. The interviewer would probe and ask further questions when interesting topics emerged. The second section lists the interview questions used when we spoke to the representative for GitHub.

\section{Instructor Interviews}
\begin{itemize}
    \item Do you have any experience with using Github in aspects other than education? If so, how long have you been using Github?
    \item Did you use Github in your class?
    \begin{itemize}
        \item If yes ,What course(s) did you use Github for? How many semesters you used Github for teaching?
        \item If no, why not?
    \end{itemize}
    \item What was your motivation to use GitHub in your class?
    \item What was the class size (i.e. number of students)? Was it a frontal (face-to-face) or virtual (online) course? Did the students have any technical background in your class?
    \item Did you provide GitHub training to the students?
    \item What was GitHub used for in your class (such as announcement/assignments/platform for resource)? What features of Github did you use in your class (for example, push, merge, pull request)? How were these features used to achieve your purpose? Please elaborate and give examples.
    \item What was missing for you in GitHub that would have helped you in your teaching?
    \item What did you (or your class / students) benefit by using Github? What were the strengths of GitHub that made it useful to you?
    \item What were the difficulties/challenges for using GitHub in your teaching?
    \item Do you think Github saved you time, comparing to the traditional ways you used before? If so, in what ways?
    \item What was students' attitude towards the usage of GitHub in the course? Did any student complain about it?
    \item Are you satisfied with the user interface? (1-10 or doesn't matter) If not, please tell us which part and why.
    \item Did the usage of Github encourage students participation in your class? Was student performance improved when using Github comparing to when they don't use it?
    \item Are you satisfied with the experience with Github overall? (1-10) How did you like the experience of using GitHub in your teaching?
    \item Any other comments to help Github to improve its application to education?
    \item Do you have any further plans to use GitHub in the future for your teaching?
    \item Do you see it useful in non technical classes (such as programming classes)?
    \item Would you like to add or comment anything else?
\end{itemize}

\section{Interview with GitHub Representative}
\begin{itemize}
    \item You mentioned (in the first email) that you work a lot with students and teachers on using GitHub in the education context. Can you tell more about this? How do you help them?
    \item In what scope do they use GitHub?
    \item What was their feedback? Most useful features? Biggest Hurdles?
    \item What do you think the motivation of educators to use GitHub?
    \item What other examples of usage of GitHub for the purposes of education have you seen (or heard of)?
    \item Any idea if teachers share materials/resources with each other?
    \item Do you see GitHub being useful to support learning and teaching? In non-technical aspects as well?
    \item What benefits does GitHub has, that might be useful for education?
    \item What challenges do you think people have in the context of education?
    \item How do you think these challenges can be mitigated?
    \item Do you think there is a way to resolve the `privacy/patriot act/us located servers/university policy' problem? How?
    \item How does GitHub (the company) support educators at it's current state?
    \item Does GitHub plans to extend it's support? In what way?
    \item How important is it for GitHub (the company) to support GitHub in the context of learning, education and teaching? (since it's not its main core).
    \item Does GitHub directly trains educators/students in using GitHub for learning? What resources does it uses for that? (this might be covered in the first question)
    \item Would you like to add or comment anything else? Share any insights?
\end{itemize}

\startappendix{Interview Questions - Case Study}
This appendix lists the interview questions I used in the study detailed in Chapter 5. The first section outlines the questions posed to each student, while the second and third sections list the questions asked of the teaching team.

\section{Student Interviews}
The first 7 items were asked of all students. As themes began emerging from the early interviews, I added the rest of the questions.
\begin{itemize}
    \item Have you used GitHub before? In what capacity? If not, have you used any similar version control tools?
    \item How have you used GitHub in your class so far?
    \begin{itemize}
        \item What do you think so far? Elaborate.
        \item Have you contributed to the course repository in any way? How? Why or why not?
    \end{itemize}
    \item Could you show me your project repository so far (if you've began your project)?
    \item Are you using GitHub in any other way than mandated in class? How?
    \item What are the challenges or drawbacks from using GitHub in your class so far? How do you think they can be mitigated?
    \item What are some features or things that GitHub could use to have assisted your class or group?
    \item What would you change, if anything, about how the course has used GitHub?
    \item What is your opinion on your course work being public in GitHub compared to being private in Coursespaces?
    \item What is your opinion regarding the education of using tools like Git, GitHub, and other DVCS in UVic?
    \item Have you looked at other public Git repositories as you wrote your project? Why?
    \item What are your opinions on GitHub's use as an LMS as compared to other LMS like Connex or Coursespaces?
    \item You had a lab or two where you were required to look at each other's work and make comments. What did you think of this exercise?
\end{itemize}

\section{Course Instructor Interview}
\begin{itemize}
    \item Do you believe you benefited from the use of GitHub in your class? How?
    \item Do you believe the students benefited from the use of GitHub in your class? How?
    \item What were the main challenges you encountered? If a professor you know were trying it for their class, what would your advice be?
    \item Do you think you might use GitHub or similar tools in your classes in the future? Why or why not?
    \item What features of GitHub did you feel were the most useful for a classroom environment? Any features that you wish you utilized more of that the students would use more?
    \item What improvements would you suggest, if any, for GitHub to be useful as a learning tool?
    \item How do you think the students felt overall about using GitHub in this way?
    \item How do you feel GH as a learning tool compares to more traditional LMS or other tools you've used in your classes?
\end{itemize}

\section{Lab Instructor Interviews}
\begin{itemize}
    \item What have you TA'd before this course? What tools were used then, if any?
    \item How much experience have you had with GH before the course?
    \item What was Yvonne's instructions in terms of using GitHub for the labs?
    \item What was Yvonne's instructions in terms of using GitHub for the labs?
    \item What are some of the difficulties you experienced with using GitHub, if any?
    \item Do you look at student projects? How? On their repositories?
    \item What do you think of GitHub as an LMS? What are the advantages and disadvantages compared to traditional LMS like Connex and Coursespaces?
    \item Anything you think you would have wanted to do differently?
\end{itemize}

\startappendix{Questionnaire \& Survey - Student Perspective}
This appendix lists the questions used in the early questionnaire and in the validation survey for the study highlighted in Chapter 5. The first section lists the questions from the Questionnaire, given to the students soon after the course began. These questions were mostly open-ended questions. The second section lists the questions from the Survey, which served to validate or invalidate my findings. These questions were mostly Likert Scale type questions, where 1 indicated that they strongly disagreed with the question and 5 indicated that they strongly agreed.

\section{Questionnaire}
\begin{itemize}
    \item How would you rate your familiarity with Distributed Version Control (such as Git, Mercurial, etc.)? [1-5]
    \item In which contexts have you used GitHub before? [Work, Class, Group Projects, Personal Projects]
    \item How would you rate your familiarity with GitHub? [1-5]
    \item From what you know of GitHub, what are your thoughts on how its use might benefit your education?
    \item From what you know of GitHub, what are your thoughts on the challenges of using it in your education?
    \item What is your favorite learning management tool? [Connex, Coursespaces, Moodle, Other]
    \item What do you think are some of the most important features of that learning management tool? And why do you feel these features are important?
    \item If you have used GitHub in a class before, please discuss your experiences.
    \item Is there anything else you'd like to comment on?
\end{itemize}

\section{Validation Survey}
\begin{itemize}
    \item What level of study are you in? [Second, Third, Fourth, Fifth+, Graduate Student]
    \item What program are you in? [Computer Science, Software Engineering, Other]
    \item Did we already interview you about this topic? Or did you sign up to be interviewed in the next two weeks? [Yes, No]
    \item If we haven't yet interviewed you and you are up for a 20-30 minute interview about this topic, please enter your email address so we can learn more about your responses!
    \item BEFORE this class, I have used Git \& GitHub in the following contexts: [Class, Group Projects, Work, Projects]
    \item BEFORE this course, this was my level of familiarity with GitHub: [1-5]
    \item Currently, this is my level of familiarity with GitHub: [1-5]
    \item I enjoyed how GitHub overall how GitHub was used in our class, labs, and projects. [1-5]
    \item I will continue using GitHub for group projects. [1-5]
    \item I will continue using GitHub as a code repository for individual work. [1-5]
    \item GitHub was useful for organizing our group projects. [1-5]
    \item Being able to trace or account for my own or others' activities was useful to me. [1-5]
    \item The ability to easily view others' projects and comment on them made me feel more involved in the class [1-5]
    \item Others commenting on my work or answering my questions was useful for my projects. [1-5]
    \item The option of being able to make pull requests for the course is appealing to me. [1-5]
    \item I found using Git \& GitHub for this course easy to learn and use. [1-5]
    \item Having all my school work and projects in one place is beneficial for my career. [1-5]
    \item Using GitHub in this class helped me learn a lot more about Git \& GitHub. [1-5]
    \item Having learned a distributed version control system such as Git is beneficial for my career. [1-5]
    \item I don't like the idea of my work for school being out in the public. [1-5]
    \item I don't like the idea of my comments for my classes being out in the public. [1-5]
    \item I disliked that we had two systems set up for this course. [1-5]
    \item I felt there wasn't enough collaboration in this class or contribution from the students to justify using GitHub for the course. [1-5]
    \item I had difficulties learning how to use Git \& GitHub for this class. [1-5]
    \item I feel this class should have offered a tutorial in using Git \& GitHub right at the beginning. [1-5]
    \item I feel that Git, GitHub, and other Distributed Version Control Systems should play a bigger role in our education at UVic. [1-5]
    \item I feel that GitHub features are sufficient such that it can serve as a standalone Learning Management System [1-5]
    \item I think the GitHub file organization is much cleaner and easier to read than other LMS I've used. [1-5]
    \item These are some features that I think GitHub needs so that it can better support education.
    \item The workflow in using Git \& GitHub in this class was sufficient. [1-5]
    \item I would like to see more classes in the future using GitHub as the primary class platform. [1-5]
    \item Do you have any further suggestions on improving the workflow of using GitHub to better support learning in your courses?
    \item Do you have any suggestions on how GitHub itself can be improved so it can better support learning in your courses?
    \item Is there anything else you'd like to add or comment on?
    \item If you would like to be notified of our results, please enter your email address below
\end{itemize}
